Before starting, I would like to take this opportunity to express my gratitude. I would like to thank all the people that are close to me for their continuous support and motivation to study electronic and computer systems.

%Please write all your preface text here. If you do so, don't forget to thank your supervisor, other committee members, your family, colleagues etc.\ etc. 
I want to thank my mom who is always there for me and my brother who always helps me in many ways and the people close to me for their continuous support and motivation to study electronic and computer systems.

Furthermore, I would also like to thank the leading expert on computer architectures, professor dr. Henk Corporaal for guidance and introducing me to the field. %introduced %intrigued TODO: Q: intrigued or introduced..? both are very ince 									 me to study computer architectures
Many thanks also go to my supervisor from Radbound university in Nijmegen, dr. ir. Roel Jordans for guidance and help on the compiler, and for introducing me to compilers.

Also, many thanks go to ir. Luc Waeijen who helpt me in many of the technical aspects of this master project and to dr. ir. Lech Jozwiak, who always helps a lot by providing extremely useful feedback.

More thanks go to dr. ir. Pieter Cuijpers for joining my committee, and I would like to thank Liu Zhenyuan for his initial implementation of the compiler and Boyan Liang for his view on the hardware side of this project.

Finally, I would like to conclude the thanks with a quote, ``Tell me and I forget, Teach me and I remember. Involve me and I learn.'' - by Benjamin Franklin.

%TODO: move henk first, parents and bro last.
%TODO: thank to Liu, for his initial implementation of compiler

%I would like to end this thanks with a quote, ``Tell me and I forget, Teach me and I remember. Involve me and I learn." - Benjamin Franklin.

%thanks to electronic systems group, my family and brother that guided and supported me throughout my entire study. blablabla
